---
title: "USUniversities.Rmd"
author: "Varun Parameswaran, Stanley Urbanek"
date: "October 2, 2015"
output: html_document
---

#University Demographics and Debt by State

####This is project 2 for Dr. Cannata's CS 329 Data Visualization class.

####This is how my RStudio is configured:

```{r}
sessionInfo()
```

### Introduction

We wished to know about the demographics of universities around the United States. We located a data set that was a list of ~8000 US universities with demographic make up, SAT Scores, initial income post graudation, and percentage taking a financial loan. We will run 3 experiments to draw conclusions about US Universities.

We operated on a CSV file, uploaded to the Oracle database, in R. Below, we initialize our system. 

```{r}
source("../01 Data/Initialize.R", echo = TRUE)
```

Here is our R_ETL file's operations, which allow us to "clean" the data in RStudio and export it to SQL.

```{r}
source("../01 Data/R_ETL.R", echo = TRUE)
```

We used the above SQL statement to create the College Grad table in SQL.

Finally, here is our retrieval file, where we pull the data back from SQL and display the "head" and "summary" properties of our data.

```{r}
source("../01 Data/Retrieval.R", echo = TRUE)
```

### Hypothesis 1: Average University Debt by State

##### Introduction

What we want to know is, what is the most indebting state/region of the country to attend college? In this data, we can extract generalizations about the quality of colleges in the region (more indebting ~> more expensive ~> higher quality facilities, faculty), and about the economic situation of the area (more expensive ~> better/worse fiscally, or far more people in the area). What we expect, is the North East, and California, to be the most expensive, due to concentration of higher tier schools, such as Harvard, MIT, and CalTech. We expect the Pacific Island region, and midwest, area to be the least expensive.

##### Table and Visualization

```{r}
source("../02 Data Wrangling/Workflow1.R", echo = TRUE)
```


##### Analysis
From this data, we can see that, on average, South Dakota, Vermont and Nebraska are the most indebting states, while Wyoming, New Jersey, and Hawaii are the least. This does not support our hypothesis. This is interesting because this tells us that South Dakota and Nebraska, while more sparsely populated, have presumably a highest quality colleges/high density of affluent people. This indicates that they might be considered a destination college location. Conversely, this is interesting because Hawaii is one of the least indebting, perhaps hinging on the quality they provide/demographic they service, or some lower cost of living. 


###Experiment 2: 

##### Introduction

Our second experiment expands on the first. We want to know how the average percentage of students needing loans correlates with the average debt upon graduation, and analyze this correlation by regions. This is meaningful because it is a better indicator of the financial aid and cost of attendance at these states. For example, if a particular state has a relatively high average debt upon graduation compared to a relatively low percentage of students needing loans, this suggests that the cost of attendance in that state may be higher (increasing average debt upon graduation) but that more students would have financial aid at these schools (decreasing the percentage of students needing loans). Our hypothesis is that states in the Northeast and North Central regions will have higher debts coupled with lower percentage seeking loans, while states in the South and West will have the opposite.

##### Data and Visualizations

```{r}
source("../02 Data Wrangling/Workflow2.R", echo = TRUE)
```

##### Conclusions

The data wrangling and visualization revealed that our hypothesis was mostly correct. There is a fairly strong correlation between the percentage of students needing loans and average debt upon graduation; however, as we predicted, states in the northeast and north central are skewed toward having higher debts and having less percent of students needing loans. This makes sense because the cost of living in New York, Massachusetts, Pennsylvania, etc. is higher so these colleges would be more expensive. However, southern colleges required higher debts than we thought they would. Perhaps this is because the population there is generally not as affluent as in the northeast, so students need to borrow more money there.

###Experiment 3: Examining success in high school and post-graduation, with race

##### Introduction

We wish to know how success in high school, and after college graduation, relates to race. We quantify success in high school as an average SAT score, and success after college graduation as initial income. Based on presumed US stereotypes of success by race, we would hypothesize that whites blanket and establish a heavy backgroup and average, Asians perform above average in both, while Blacks and Hispanics score low-medium average for both. Just to clarify - for this visualization, we calculated the majority race in each college and assigned that college's data point the appropriate color.

##### Table and Visualizations


```{r}
source("../02 Data Wrangling/Workflow3.R", echo = TRUE)
```

##### Conclusions
The data wrangling and visualization revealed that White demographic establishes the average, Asians perform above average in both, while Blacks and Hispanics score low-medium average for both. This supports our hypothesis. This is interesting because it demonstrates that there may be some truth behind the racial stereotypes in the US. At least, it indicates that there may be some grounding in a truth.


